Summer 2025 ToDo

### 2025-09-05 Relations/Paper/Definitions.tex
Changes made on meeting 8th July
Symbol changes reflected in todo list

% \subsection{Confluence (\texttt{Properties.agda})}
\begin{definition} Let $a \in A$.
    \begin{description}
        \item[\ule{$a \in \WCR_R$}] \emph{$a$ is weakly Church-Rosser} (or \emph{weakly confluent}) if for all $b,c \in A$, \[c \bstep a \rstep b \implies \exists d.\, c \mstep d \bmstep b\]
        \item[\ule{$a \in \CR_R$}] \emph{$a$ is Church-Rosser} (or \emph{confluent}) if for all $b,c \in A$, \[c \bmstep a \mstep b \implies \exists d.\, c \mstep d \bmstep b\]
        \item[\ule{$a \in \CRs_R$}] \emph{$a$ is subcommutative} if for all $b, c \in A$,
        \[c \bmstep a \mstep b \implies \exists d. \, c \rrstep d \brstep b\]
        \item[\ule{$a \in \NPe_R$}] \emph{$a$ has the conversion normal form property} if for all $b \in \NF_R$, \[ a \estep b \implies a\mstep b\]
        \item[\ule{$a \in \NP_R$}] \emph{$a$ has the normal form property} if for all $b \in \NF_R$,
        \[c \bmstep a \mstep b \implies c \mstep b\]
        \item[\ule{$a \in \UN_R$}] \emph{$a$ has the unique normal form property} if for all $a, b \in \NF_R$, \[a \estep b \implies a \equiv b\]
        \item[\ule{$a \in \UNto_R$}] \emph{$a$ has the unique normal form property with respect to reduction} if for all $b, c \in \NF_R$,
        \[c \bmstep a \mstep b  \implies b \equiv c\]
    \end{description}
\end{definition}
% \begin{definition} Let $a \in A$.
%     \begin{description}
%         \item[\ule{$a \in \WCR_R$}] \emph{$a$ is weakly Church-Rosser (or weakly confluent)} if $c \bstep a \rstep b \implies c \mstep d \bmstep b$.
%         \item[\ule{$a \in \CR_R$}] \emph{$a$ is Church-Rosser (or confluent)} if $c \bmstep a \mstep b \implies c \mstep d \bmstep b$.
%         \item[\ule{$a \in \CRs_R$}] \emph{$a$ is subcommutative} if $c \bmstep a \mstep b \implies c \rrstep d \brstep b$.
%         \item[\ule{$a \in \NPe_R$}] \emph{$a$ has the conversion normal form property} if $a \estep b \implies a\mstep b$ for any $b \in \NF$.
%         \item[\ule{$a \in \NP_R$}] \emph{$a$ has the normal form property} if $c \bmstep a \mstep b \implies c \mstep b$ for any $b \in \NF$.
%         \item[\ule{$a \in \UN_R$}] \emph{$a$ has the unique normal form property} if $a \estep b \implies a \equiv b$  for any $a, b \in \NF$.
%         \item[\ule{$a \in \UNto_R$}] \emph{$a$ has the unique normal form property with respect to reduction} if $c \bmstep a \mstep b  \implies b \equiv c$  for any $b, c \in \NF$.
%     \end{description}
% \end{definition}

@@

In \terese, $\mathrm{\,Ind}$ is used to denote $\gBP$.

>>>>

% \subsection{Confluence (\texttt{Properties.agda})}
\begin{definition} Let $a \in A$.
    \begin{description}
        \item[\ule{$a \in \WCR_R$}] \emph{$a$ is weakly Church-Rosser} (or \emph{weakly confluent}) if for all $b,c \in A$, \[c \bstep a \rstep b \implies \exists d.\, c \mstep d \bmstep b\]
        \item[\ule{$a \in \CR_R$}] \emph{$a$ is Church-Rosser} (or \emph{confluent}) if for all $b,c \in A$, \[c \bmstep a \mstep b \implies \exists d.\, c \mstep d \bmstep b\]
        \item[\ule{$a \in \CRs_R$}] \emph{$a$ is subcommutative} if for all $b, c \in A$,
        \[c \bmstep a \mstep b \implies \exists d. \, c \rrstep d \brstep b\]
        \item[\ule{$a \in \NPe_R$}] \emph{$a$ has the conversion normal form property} if for all $b \in \NF_R$, \[ a \estep b \implies a\mstep b\]
        \item[\ule{$a \in \NP_R$}] \emph{$a$ has the reduction normal form property} if for all $b \in \NF_R$,
        \[c \bmstep a \mstep b \implies c \mstep b\]
        \item[\ule{$a \in \UN_R$}] \emph{$a$ has the conversion unique normal form property} if for all $a, b \in \NF_R$, \[a \estep b \implies a \equiv b\]
        \item[\ule{$a \in \UNto_R$}] \emph{$a$ has the reduction unique normal form property} if for all $b, c \in \NF_R$,
        \[c \bmstep a \mstep b  \implies b \equiv c\]
    \end{description}
\end{definition}
% \begin{definition} Let $a \in A$.
%     \begin{description}
%         \item[\ule{$a \in \WCR_R$}] \emph{$a$ is weakly Church-Rosser (or weakly confluent)} if $c \bstep a \rstep b \implies c \mstep d \bmstep b$.
%         \item[\ule{$a \in \CR_R$}] \emph{$a$ is Church-Rosser (or confluent)} if $c \bmstep a \mstep b \implies c \mstep d \bmstep b$.
%         \item[\ule{$a \in \CRs_R$}] \emph{$a$ is subcommutative} if $c \bmstep a \mstep b \implies c \rrstep d \brstep b$.
%         \item[\ule{$a \in \NPe_R$}] \emph{$a$ has the conversion normal form property} if $a \estep b \implies a\mstep b$ for any $b \in \NF$.
%         \item[\ule{$a \in \NP_R$}] \emph{$a$ has the normal form property} if $c \bmstep a \mstep b \implies c \mstep b$ for any $b \in \NF$.
%         \item[\ule{$a \in \UN_R$}] \emph{$a$ has the unique normal form property} if $a \estep b \implies a \equiv b$  for any $a, b \in \NF$.
%         \item[\ule{$a \in \UNto_R$}] \emph{$a$ has the unique normal form property with respect to reduction} if $c \bmstep a \mstep b  \implies b \equiv c$  for any $b, c \in \NF$.
%     \end{description}
% \end{definition}

@@

The property $\gBP$ is equivalent to the property called $\mathrm{\,Ind}$ in \terese. (Specifically, $\mathrm{\,Ind}$ asserts the exists of a bound for finite sequences as well, but that is automatic, since the last term of a finite reduction sequence is a bound for that sequence.)

### 2025-09-05 Relations/Paper/Summer_2025_TODO.tex
Symbol changes reflected in todo list

\section*{20th August todos}
\begin{todolist}
  \item (Secondary) Find consistency with NP and UN with respect to reduction and conversion. Should the default be by reduction for both. Should we be explicit with $\to$ and = in each case.  \sacomment{Main issue here is that we'll either introduce an inconsistency with our code or we'll have a hard to read symbol. We can do $\to$ as super or sub-script in Agda. If we're happy with $NP\to$ in agda but using a different script in the paper then I think we should go ahead and be explicit for both UN and NP as to whether it is WRT conversion or reduction. We already have this inconsistency in the code with $UN\to$} 
  \item Carry on modifying the names of wfXXX- to wfXXXnotnot with prefix underscore in agda and in latex. \sacomment{Done with the exception of WFseq-. This needs a bit more thought. Similarly for the relation between WFcor and WFminCor.}
  \item Make a decision regarding WFcor. We use it to imply WFacc using AccCor. But otherwise it is sort of orphaned off. Does it help contribute to the overall story? \sacomment{We've now related it to WFmin by requiring that correductive predicates have double negation elimination. Is this too strong a requirement? Similarly we have linked WFacc to WFcor but this proof requires the same property as well as limiting ourselves to accessible predicates. Again, is this too much?}
  \item Get the classical hypothesis and relations organized and see how they're used in the implications (organize implications by classical requirement. In the implications file we can import the classical implications. Basically, refactor the code.) \sacomment{Done but could be further refactored.}
  \item Make a cheat sheet for the next meeting regarding to talk about coreductive. How it relates on the weak side and on the strong side. SEQ seq+- , how to present these. Try and clarify what the issues are here and what exactly are the differences between them in terms of definition and implication. Could one be removed and the overall story stay the same? \sacomment{Regarding seq+ (and seq+-) we currently only show that seq+ implies seq and seq+- implies seq- so could be removed without changing any of our other implictions.}
\end{todolist}

>>>>

\section*{20th August todos}
\begin{todolist}
  \item [\done] (Secondary) Find consistency with NP and UN with respect to reduction and conversion. Should the default be by reduction for both. Should we be explicit with $\to$ and = in each case.  \sacomment{Main issue here is that we'll either introduce an inconsistency with our code or we'll have a hard to read symbol. We can do $\to$ as super or sub-script in Agda. If we're happy with $NP\to$ in agda but using a different script in the paper then I think we should go ahead and be explicit for both UN and NP as to whether it is WRT conversion or reduction. We already have this inconsistency in the code with $UN\to$} \apcomment{In the agda code let's not use arrow and have arrow implied as default but be explicit with '=' sign.}
  \item Carry on modifying the names of wfXXX- to wfXXXnotnot with prefix underscore in agda and in latex. \sacomment{Done with the exception of WFseq-. This needs a bit more thought. Similarly for the relation between WFcor and WFminCor.}
  \item Make a decision regarding WFcor. We use it to imply WFacc using AccCor. But otherwise it is sort of orphaned off. Does it help contribute to the overall story? \sacomment{We've now related it to WFmin by requiring that correductive predicates have double negation elimination. Is this too strong a requirement? Similarly we have linked WFacc to WFcor but this proof requires the same property as well as limiting ourselves to accessible predicates. Again, is this too much?}
  \item Get the classical hypothesis and relations organized and see how they're used in the implications (organize implications by classical requirement. In the implications file we can import the classical implications. Basically, refactor the code.) \sacomment{Done but could be further refactored.}
  \item Make a cheat sheet for the next meeting regarding to talk about coreductive. How it relates on the weak side and on the strong side. SEQ seq+- , how to present these. Try and clarify what the issues are here and what exactly are the differences between them in terms of definition and implication. Could one be removed and the overall story stay the same? \sacomment{Regarding seq+ (and seq+-) we currently only show that seq+ implies seq and seq+- implies seq- so could be removed without changing any of our other implictions.} 
\end{todolist}

### 2025-09-05 Relations/Paper/Well-foundedness.tex
Symbol changes reflected in todo list

\begin{definition}
  Let $R \subseteq A \times A$. We write $Rxy$ when $(x,y) \in R$.
  \begin{enumerate}
    \item $R$ is \emph{accessibly well-founded} ($\WFacc$) if every element is accessible:
      \[
        \forall x \in A.\; x \text{ is accessible}.
      \]
    \item $R$ is \emph{inductively well-founded} ($\WFind$) if every inductive $P \subseteq A$ is universally true:
      \[
        \forall P \subseteq A.\;\text{$P$ is inductive} \Rightarrow \forall x \in A.\; x \in P.
      \]
    \item $R$ is \emph{sequentially well-founded} ($\WFseq$) if every sequence contains an index at which it is not decreasing:
      \[
        \forall s : \nat \to A \exists k : \nat.\, \lnot \, R \,(s\,k)\,(s\,(k+1)).
      \]
    \item  $R$ is \emph{minimally well-founded} ($\WFmin$) if every nonempty subset has a $P$-minimal element:
      \[
        \forall P \subseteq A.\; P \neq \varnothing \Rightarrow
        \exists x \in P.\; \text{$x$ is $P$-minimal}.
      \]
    \item  $R$ is \emph{minimally well-founded for $\lnot \lnot$-closed predicates} ($\WFminDNE$) if every nonempty $\lnot \lnot$-closed subset contains a $P$-minimal element:
      \[
        \forall P \subseteq A.\; P \neq \varnothing \Rightarrow \left(\ol{\ol P} \subseteq P\right) \Rightarrow
        \exists x \in P.\; \text{$x$ is $P$-minimal}.
      \]
    \item $R$ is \emph{coreductively well-founded} ($\WFcor$) if every coreductive $P \subseteq A$ is universally true:
      \[
        \forall P \subseteq A.\;(\text{$P$ is coreductive}) \Rightarrow \forall x \in A.\; x \in P.
      \]

@@

\begin{definition}
  Let $R \subseteq A \times A$. We write $Rxy$ when $(x,y) \in R$.
  \begin{enumerate}
    \item $R$ is \emph{weakly accessibly well-founded} ($\WFaccm$) if accessibility holds up to double negation:
      \[
        \forall x \in A.\; \lnot \lnot  (x \text{ is accessible}).
      \]
    \item $R$ is \emph{weakly inductively well-founded} ($\WFindm$) if every inductive predicate holds for all elements up to double negation:
      \[
        \forall P \subseteq A.\;(\text{$P$ is inductive}) \Rightarrow \forall x \in A.\; \lnot \lnot (x \in P).
      \]
    \item $R$ is \emph{weakly sequentially well-founded} ($\WFseqm$) if no sequence is $R$-decreasing:
      \[
        \forall s:\nat \to A.\; \lnot \text{($s$ is $R$-decreasing)}.
      \]
    \item  $R$ is \emph{weakly minimally well-founded} ($\WFminm$) if every nonempty subset contains a $\lnot \lnot P$-minimal element:
      \[
        \forall P \subseteq A.\; P \neq \varnothing \Rightarrow \lnot \lnot \;
        (\exists x \in P.\; \text{$x$ is $P$-minimal}).
      \]
    \item  $R$ is \emph{weakly minimally well-founded under double negation elimination} ($\WFminDNEm$) if every nonempty $\lnot \lnot$-closed subset $\lnot \lnot$ contains a $P$-minimal element:
      \[
        \forall P \subseteq A.\; P \neq \varnothing \Rightarrow \left(\ol{\ol P} \subseteq P\right) \Rightarrow \lnot \lnot \;
        (\exists x \in P.\; \text{$x$ is $P$-minimal}).
      \]      
  \end{enumerate}
\end{definition}

@@

\draw[arrow, bend right = 90] (WFacc) to node[pos = 0.75,left, text = red] {Rdec} (WFseq);
\draw[arrow, bend right] (WFaccM) to (WFminDNEM);
\draw[arrow, bend right = 90] (WFseqM) to node[right, text=red] {accCor} (WFaccM);  
% \draw[arrow, bend left = 55] (WFaccM) to (WFminM);

@@

As this discussion illustrates, the network of relationships between the
constructive notions of well-foundedness is quite rich.  The purpose of
this section is to present the basic picture of these implications,
and elaborate on some of them.  Our key observations are summarized
in the following list.
\begin{enumerate}
  \item The implication from the inductive to the sequential definition
  requires the underlying relation to be decidable. This assumption is no longer
  necessary on the $\lnot\lnot$-translated side.
  \item WFmin is a very strong assumption and implies WFseq unconditionally.
  The more reasonable WFminDNE requires Markov's Principle
  (and decidability of $R$) to yield the same consequence.
  \item On the $\lnot\lnot$-side, WFmin- and WFminDNE- are equivalent,
  and imply WFacc- via WFseq-, requiring {\accCor} assumption in the final inference.
  \item At the same time, if the relation is finitely branching, then WFminDNE-
  implies WFacc- directly.
  \item Finite branchingness together with (weak) decidability of minimality
  implies \accCor.
  % \apcomment{The previous observation suggests the question of whether isWFseq- might
  % imply isWFminDNE- via decidability of minimal elements.\\
  % However, in trying to prove this implication, the hypothesis asserting non-existence of
  %  a minimal element of some given non-empty $\lnot\lnot$-closed predicate
  %  automatically entails that any particular element of interest is not normal.\\
  %  Unless it's not satisfying $P$? \emph{We need to decide whether it is P-minimal as well.}}
  \item Assuming \accCor, isWFacc- can also be derived from a weaker condition, isWFminCor(-).
  (Just like isWFminDNE is a natural weakening of isWFmin that is sufficient to recover isWFacc
  in the presence of {\accDNE}.)
  It is implied by isWFseq-, but does not appear to imply anything on its own.

  \apcomment{(This seems easy: The converse implication isWFminCor to isWFseq- requires the sequence in question to
  have decidable image, leading to a further weakening of isWFseq-.)}
  \item Since isWFminDNE- implies isWFseq-, the former also implies isWFminCor-.
  But the converse fails: If a subset/predicate $P$ is $\lnot\lnot$-closed,
  we cannot conclude it is coreductive.

>>>>

\begin{definition}
  Let $R \subseteq A \times A$. We write $Rxy$ when $(x,y) \in R$.
  \begin{enumerate}
    \item $R$ is \emph{accessibly well-founded} ($\WFacc$) if every element is accessible:
      \[
        \forall x \in A.\; x \text{ is accessible}.
      \]
    \item $R$ is \emph{inductively well-founded} ($\WFind$) if every inductive $P \subseteq A$ is universally true:
      \[
        \forall P \subseteq A.\;\text{$P$ is inductive} \Rightarrow \forall x \in A.\; x \in P.
      \]
    \item $R$ is \emph{sequentially well-founded} ($\WFseq$) if every sequence contains an index at which it is not decreasing:
      \[
        \forall s : \nat \to A \exists k : \nat.\, \lnot \, R \,(s\,k)\,(s\,(k+1)).
      \]
    \item  $R$ is \emph{well-founded with regard to minimality} ($\WFmin$) if every nonempty subset has a $P$-minimal element:
      \[
        \forall P \subseteq A.\; P \neq \varnothing \Rightarrow
        \exists x \in P.\; \text{$x$ is $P$-minimal}.
      \]
    \item  $R$ is \emph{well-founded with regard to minimality for $\lnot \lnot$-closed predicates} ($\WFminDNE$) if every nonempty $\lnot \lnot$-closed subset contains a $P$-minimal element:
      \[
        \forall P \subseteq A.\; P \neq \varnothing \Rightarrow \left(\ol{\ol P} \subseteq P\right) \Rightarrow
        \exists x \in P.\; \text{$x$ is $P$-minimal}.
      \]
    \item $R$ is \emph{coreductively well-founded} ($\WFcor$) if every coreductive $P \subseteq A$ is universally true:
      \[
        \forall P \subseteq A.\;(\text{$P$ is coreductive}) \Rightarrow \forall x \in A.\; x \in P.
      \]

@@

\begin{definition}
  Let $R \subseteq A \times A$. We write $Rxy$ when $(x,y) \in R$.
  \begin{enumerate}
    \item $R$ is \emph{weakly accessibly well-founded} ($\WFaccm$) if accessibility holds up to double negation:
      \[
        \forall x \in A.\; \lnot \lnot  (x \text{ is accessible}).
      \]
    \item $R$ is \emph{weakly inductively well-founded} ($\WFindm$) if every inductive predicate holds for all elements up to double negation:
      \[
        \forall P \subseteq A.\;(\text{$P$ is inductive}) \Rightarrow \forall x \in A.\; \lnot \lnot (x \in P).
      \]
    \item $R$ is \emph{weakly sequentially well-founded} ($\WFseqm$) if no sequence is $R$-decreasing:
      \[
        \forall s:\nat \to A.\; \lnot \text{($s$ is $R$-decreasing)}.
      \]
    \item  $R$ is \emph{weakly well-founded with regard to a minimal element} ($\WFminm$) if for every nonempty subset there $\lnot \lnot$-exists a $P$-minimal element:
      \[
        \forall P \subseteq A.\; P \neq \varnothing \Rightarrow \lnot \lnot \;
        (\exists x \in P.\; \text{$x$ is $P$-minimal}).
      \]
    \item  $R$ is \emph{weakly well-founded with regard to a minimal element under double negation elimination} ($\WFminDNEm$) if every nonempty $\lnot \lnot$-closed subset $\lnot \lnot$-contains a $P$-minimal element:
      \[
        \forall P \subseteq A.\; P \neq \varnothing \Rightarrow \left(\ol{\ol P} \subseteq P\right) \Rightarrow \lnot \lnot \;
        (\exists x \in P.\; \text{$x$ is $P$-minimal}).
      \]      
    \item [need to define wfmincor]
  \end{enumerate} 
\end{definition}

@@

\draw[arrow, bend right = 90] (WFacc) to node[pos = 0.75,left, text = red] {Rdec} (WFseq);
\draw[arrow, bend right] (WFaccM) to (WFminDNEM);
\draw[arrow, bend right = 90] (WFminCor) to node[right, text=red] {accCor} (WFaccM);  
% \draw[arrow, bend left = 55] (WFaccM) to (WFminM);

@@

As this discussion illustrates, the network of relationships between the
constructive notions of well-foundedness is quite rich.  The purpose of
this section is to present the basic picture of these implications,
and elaborate on some of them.  Our key observations are summarized
in the following list.
\begin{enumerate}
  \item The implication from the inductive to the sequential definition
  requires the underlying relation to be decidable. This assumption is no longer
  necessary on the $\lnot\lnot$-translated side.
  \item WFmin is a very strong assumption and implies WFseq unconditionally.
  The more reasonable WFminDNE requires Markov's Principle
  (and decidability of $R$) to yield the same consequence.
  \item On the $\lnot\lnot$-side, WFmin- and WFminDNE- are equivalent,
  and imply WFacc- via WFseq-, requiring {\accCor} assumption in the final inference.
  \item At the same time, if the relation is finitely branching, then WFminDNE-
  implies WFacc- directly.
  \item Finite branchingness together with (weak) decidability of minimality
  implies \accCor.
  % \apcomment{The previous observation suggests the question of whether isWFseq- might
  % imply isWFminDNE- via decidability of minimal elements.\\
  % However, in trying to prove this implication, the hypothesis asserting non-existence of
  %  a minimal element of some given non-empty $\lnot\lnot$-closed predicate
  %  automatically entails that any particular element of interest is not normal.\\
  %  Unless it's not satisfying $P$? \emph{We need to decide whether it is P-minimal as well.}}
  \item Assuming \accCor, $\isWFacc$ ($\isWFaccm$) can also be derived from a weaker condition, $\isWFmin$ ($\isWFminm$).
  (Just like $\isWFminDNE$ is a natural weakening of $\isWFmin$ that is sufficient to recover $\isWFacc$ in the presence of {\accDNE}.)
  It is implied by $\isWFseqm$, but does not appear to imply anything on its own.

  % \apcomment{(This seems easy: The converse implication isWFminCor to isWFseq- requires the sequence in question to
  % have decidable image, leading to a further weakening of isWFseq-.)}
  \item Since isWFminDNE- implies isWFseq-, the former also implies isWFminCor-.
  But the converse fails: If a subset/predicate $P$ is $\lnot\lnot$-closed,
  we cannot conclude it is coreductive.
